Title of Thesis (title case, double spaced, no more than 240 characters. No bold font or symbols, except Greek) \\ \bigskip Thesis Title Second Line

### 2017-07-22 case_study/case_study.tex
adds the case study paper

\documentclass[conference]{IEEEtran}

\ifCLASSINFOpdf
\else
\fi
\usepackage{url}
\usepackage[hypcap]{caption}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{eurosym}
\usepackage{multirow}
\usepackage{subfigure}

%\usepackage{caption}
%\usepackage{subcaption}

\usetikzlibrary{backgrounds}
\pgfplotsset{compat=1.5}
\usepgfplotslibrary{dateplot}

\begin{document}
\title{A Case Study of Web API Evolution}

\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, Frank Maurer
\IEEEauthorblockA{Department of Computer Science\\
University of Calgary,
Calgary, Alberta T2N 1N4\\ Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca }
}}

\maketitle

\begin{abstract}
When applications are integrated using web APIs, changes on a web API may break the dependent applications. This problem exists because old versions of the APIs may no longer be supported, a lack of adequate documentation to upgrade to a newer version, and insufficient communication of changes. In this paper we conducted a case study of evolving Web APIs to investigate what changes are made between versions and how the changes are documented and communicated to the API users. The findings are a list of recommendations for practitioners and researchers based on API change profiles, versioning, documentation and communication approaches that are observed in practice. This study will help inform developers of evolving Web APIs to make decision about versioning, documentation and communication methods.
\end{abstract}

\begin{IEEEkeywords}
Case Study; RESTful, SOAP, Web API Evolution; WSDL
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

@@

%{|p{0.2\linewidth}|p{0.3\linewidth}|p{0.2\line width}|p{0.2\linewidth}|}
%& &  \multicolumn{4}{ c |}{\textbf{Categories}} & & \\ \cline{3-6}
%\multirow{2}{*}{\textbf{Web API}}  & \multirow{2}{*}{\textbf{URL}} &\textbf{Industry} & \textbf{Popularity} &  \textbf{Maturity} & \textbf{Code}& \multirow{2}{*}{\\textbf{Rel.}} & \multirow{2}{*}{\\textbf{Dates}}\\

\begin{table*}[!ht]
\caption{Case Study - Representative Sample of Web APIs.}
\label{web_api_grid}
\begin{tabular} {| l | l  | l  | l | l | l | l  | l |}
\hline
\multirow{2}{*}{\textbf{Web API}} &
\multirow{2}{*}{\textbf{URL}} &
\multicolumn{4}{ c |}{\textbf{Categories}} &
\textbf{Num of} &
\multirow{2}{*}{\textbf{Dates}}\\    \cline{3-6}
  & & Industry & Popularity &  Maturity & Code& \textbf{Releases}& \\
\hline
Facebook Platform API & \tiny{\url{https://developers.facebook.com/docs/apps/versions}} & Social & High & Est & Closed & 11 & 01/2013-07/2014\\
\hline
Twitter REST API & \tiny{\url{https://dev.twitter.com/docs/api}} & Social & High & Est & Closed & 2 & Not Available\\
\hline
Wordpress REST API & \tiny{\url{http://wp-api.org/}} & Business & Growing &  New & Open & 8 & 08/2013-06/2014\\
\hline
Salesforce \& Chatter REST API & \tiny{\url{https://www.salesforce.com/us/developer/docs/api_rest/}} & Business & Popular & Est & Closed & 4 & 01/2013-08/2014\\
\hline
Google Calendar API & \tiny{\url{https://developers.google.com/google-apps/calendar/}} & Business & Popular & Est & Closed & 2 & Not Available\\
\hline
Stripe Payment Processor API & \tiny{\url{https://stripe.com/docs/api}} & Business & Growing & New & Closed & 30 &09/2011-10/2014 \\
\hline
Github API & \tiny{\url{https://developer.github.com/}} & Productivity & Growing & New & Closed  & 59 &09/2012-10/2014\\
\hline
Google Maps API & \tiny{\url{https://developers.google.com/maps/web/}} & Productivity & Popular & Est & Closed & 102 &01/2011-09/2014\\
\hline
OpenStreetMap API &\tiny{\url{http://wiki.openstreetmap.org/wiki/API}} & Productivity & Growing & New & Open & 1 & Not Available\\
\hline
\end{tabular}
\end{table*}

@@

\subsection{Selection Strategy} % (fold)
\label{sub:selection_strategy}

@@

\begin{table*}[!tbh]
\caption{List of Codes}
\label{coding_scheme}
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\textbf{Category} & \textbf{Description}\\
\hline
Add$<$APIElements$>$ & New API elements introduced: Adds a new resource, field, field value, endpoint or method\\
\hline
Remove$<$APIElements$>$ & API elements completely removed: Removes a resource, field, field value, endpoint or method\\
\hline
Change$<$APIElements$>$ & Existing API element is replaced by another: Replaces or renames a resource, field, field value, endpoint or method\\
\hline
ErrorConditionChange & Changes the possible error conditions: New, changed or removed error response \\
\hline
ChangeFieldDataType & Changes the data type of a field: Single item to an array or vice versa, number to strings or vice versa, required to optional or vice versa\\
\hline
BehaviorChange & Changes the outcome or fixes bug: Uses new formula for computing API return values\\
\hline
ChangeAuthentication & Changes how API clients are authenticated: New, changed or deprecated API authentication methods\\
\hline
ChangeAuthorization & Changes who can access which API elements: Add, remove or modify permissions of API user\\
\hline
\end{tabular}
\end{table*}

@@

Another example that demonstrates a behavior change can be found from the 2013-10-29 version of Stripe API \footnote{\url{https://stripe.com/docs/upgrades#2014-01-31}}:

@@

\textbf{Error condition change} - Web APIs change the error conditions that may require the users to adapt their integration. For example, in Salesforce API Version 29 the error codes were changed \footnote{\url{https://na1.salesforce.com/help/pdfs/en/salesforce_winter14_release_notes.pdf#rn_186_api_objects}}:

@@

We analyzed different approaches to identify the versions of Web APIs to understand if the version identifiers can be used to infer information about their evolution. We can categorize the Web APIs into these groups:

\begin{figure*}[!htp]
\subfigure[Facebook API Changes (11 Releases)]{
\def\dataSourceFileName{facebookx.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    ybar stacked,
    bar width=5pt,
    %title=Facebook API Changes (11 Releases),
    date coordinates in=x,
    xticklabel = \year-\month,
    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend style={legend pos=north west,font=\tiny, legend columns=4},
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:facebook-chart}
}
\subfigure[Stripe API Changes (30 Releases)]{
\def\dataSourceFileName{stripe.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=Stripe API Changes (30 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:stripe-chart}
}
\subfigure[WordPress REST API Changes (8 Releases)]{
\def\dataSourceFileName{wordpress.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=WordPress REST API Changes (8 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:wordpress-chart}
}
\subfigure[GitHub API Changes (59 Releases)]{
\def\dataSourceFileName{github.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=GitHub API Changes (59 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:github-chart}
}
\subfigure[Google Maps Web API (102 Releases)]{
\def\dataSourceFileName{gmap.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=Google Maps Web API (102 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:gmaps-chart}
}
\subfigure[Salesforce REST API (4 Releases)]{
\def\dataSourceName{Salesforce Force.com and Chatter REST}
\def\dataSourceFileName{salesforce.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=Salesforce REST API (4 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
    %label={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:salesforce-chart}
}
%% Add the legend here
\begin{center}
\subfigure{
\includegraphics[keepaspectratio, width=10cm]{legend.png}
}
\end{center}
\caption{Case Study - Change Profiles for the Web APIs.}
\label{fig:proflie}
\end{figure*}

\textbf{Numbered identifiers} - Contiguous integers are used to identify new versions by Salesforce and GitHub Web APIs. Such naming scheme clearly communicates the sequence between versions but cannot be used to infer compatibility between versions.

@@

@@

Maintaining multiple versions of the Web APIs provide more scheduling flexibility to the API clients but requires more effort from the API publisher. As a result, not all the older versions are maintained. When API clients lag far behind the latest version, migrating to the latest version becomes harder. For example, here is an excerpt of a comment from a Stripe Web API client developer \footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/#!searchin/api-discuss/version/api-discuss/x8TM-yHnhYQ/y9JJKytSiUAJ}}:

@@

%
%
%\begin{figure*}[!htp]
%\begin{subfigure}[b]{\textwidth}
%\def\dataSourceFileName{facebookx.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    ybar stacked,
%    bar width=5pt,
%    %title=Facebook API Changes (11 Releases),
%    date coordinates in=x,
%    xticklabel = \year-\month,
%    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend style={legend pos=north west,font=\tiny, legend columns=4},
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\caption{Facebook API Changes (11 Releases)}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{Stripe API Changes (30 Releases)}
%\def\dataSourceFileName{stripe.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=Stripe API Changes (30 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{WordPress REST API Changes (8 Releases)}
%\def\dataSourceFileName{wordpress.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=WordPress REST API Changes (8 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{GitHub API Changes (59 Releases)}
%\def\dataSourceFileName{github.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=GitHub API Changes (59 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{Google Maps Web API (102 Releases}
%\def\dataSourceFileName{gmap.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=Google Maps Web API (102 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{[Salesforce REST API (4 Releases)]}
%\def\dataSourceName{Salesforce Force.com and Chatter REST}
%\def\dataSourceFileName{salesforce.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=Salesforce REST API (4 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
%    %label={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=0.9\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%\ref{legend}
%\end{subfigure}
%
%\caption{Case Study - Change Profiles for the Web APIs.}
%\end{figure*}
%

%%%%%%%%%%%%%%%%%%
%
%x.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    ybar stacked,
%    bar width=5pt,
%    title=Facebook API Changes (11 Releases),
%    date coordinates in=x,
%    xticklabel = \year-\month,
%    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend style={legend pos=north west,font=\tiny, legend columns=4},
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%
%\ref{legend}
%
%\def\dataSourceFileName{stripe.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=Stripe API Changes (30 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%
%\def\dataSourceFileName{wordpress.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=WorPress REST API Changes (8 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%
%\def\dataSourceFileName{github.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=GitHub API Changes (59 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%\def\dataSourceFileName{gmap.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=Google Maps Web API (102 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%\def\dataSourceName{Salesforce Force.com and Chatter REST}

>>>>

\usetikzlibrary{backgrounds}
\pgfplotsset{compat=1.5}

\chapter{A Case Study of Web API Evolution}

%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, Frank Maurer
%\IEEEauthorblockA{Department of Computer Science\\
%University of Calgary,
%Calgary, Alberta T2N 1N4\\ Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca }
%}}

%\maketitle

\section{Abstract}
When applications are integrated using web APIs, changes on a web API may break the dependent applications. This problem exists because old versions of the APIs may no longer be supported, a lack of adequate documentation to upgrade to a newer version, and insufficient communication of changes. In this paper we conducted a case study of evolving Web APIs to investigate what changes are made between versions and how the changes are documented and communicated to the API users. The findings are a list of recommendations for practitioners and researchers based on API change profiles, versioning, documentation and communication approaches that are observed in practice. This study will help inform developers of evolving Web APIs to make decision about versioning, documentation and communication methods.

%\begin{IEEEkeywords}
%Case Study; RESTful, SOAP, Web API Evolution; WSDL
%\end{IEEEkeywords}

%\IEEEpeerreviewmaketitle

@@

@@

\subsection{Selection Strategy} % (fold)
\label{sub:selection_strategy}
\begin{table*}[!htbp]
\caption{Case Study - Representative Sample of Web APIs.}
\label{web_api_grid}
\rotatebox{-90}{
  \begin{tabular} {| p{1.5in}| p{2in}| l  | l | l | l | l  | l |}
    \hline
    \multirow{2}{*}{\textbf{Web API}} &
    \multirow{2}{*}{\textbf{URL}} &
    \multicolumn{4}{ c |}{\textbf{Categories}} &
    \textbf{Num of} &
    \multirow{2}{*}{\textbf{Dates}}\\    \cline{3-6}
    & & Industry & Popularity &  Maturity & Code& \textbf{Releases}& \\
    \hline
    Facebook Platform API & \tiny{\url{https://developers.facebook.com/docs/apps/versions}} & Social & High & Est & Closed & 11 & 01/2013-07/2014\\
    \hline
    Twitter REST API & \tiny{\url{https://dev.twitter.com/docs/api}} & Social & High & Est & Closed & 2 & Not Available\\
    \hline
    Wordpress REST API & \tiny{\url{http://wp-api.org/}} & Business & Growing &  New & Open & 8 & 08/2013-06/2014\\
    \hline
    Salesforce \& Chatter REST API & \tiny{\url{https://www.salesforce.com/us/developer/docs/api_rest/}} & Business & Popular & Est & Closed & 4 & 01/2013-08/2014\\
    \hline
    Google Calendar API & \tiny{\url{https://developers.google.com/google-apps/calendar/}} & Business & Popular & Est & Closed & 2 & Not Available\\
    \hline
    Stripe Payment Processor API & \tiny{\url{https://stripe.com/docs/api}} & Business & Growing & New & Closed & 30 &09/2011-10/2014 \\
    \hline
    Github API & \tiny{\url{https://developer.github.com/}} & Productivity & Growing & New & Closed  & 59 &09/2012-10/2014\\
    \hline
    Google Maps API & \tiny{\url{https://developers.google.com/maps/web/}} & Productivity & Popular & Est & Closed & 102 &01/2011-09/2014\\
    \hline
    OpenStreetMap API &\tiny{\url{http://wiki.openstreetmap.org/wiki/API}} & Productivity & Growing & New & Open & 1 & Not Available\\
    \hline
  \end{tabular}
}
\end{table*}

@@

\begin{table*}[!htbp]
\caption{List of Codes}
\label{coding_scheme}
\begin{tabular}{|p{0.3\linewidth}|p{0.6\linewidth}|}
\hline
\textbf{Category} & \textbf{Description}\\
\hline
Add$<$APIElements$>$ & New API elements introduced: Adds a new resource, field, field value, endpoint or method\\
\hline
Remove$<$APIElements$>$ & API elements completely removed: Removes a resource, field, field value, endpoint or method\\
\hline
Change$<$APIElements$>$ & Existing API element is replaced by another: Replaces or renames a resource, field, field value, endpoint or method\\
\hline
ErrorConditionChange & Changes the possible error conditions: New, changed or removed error response \\
\hline
ChangeFieldDataType & Changes the data type of a field: Single item to an array or vice versa, number to strings or vice versa, required to optional or vice versa\\
\hline
BehaviorChange & Changes the outcome or fixes bug: Uses new formula for computing API return values\\
\hline
ChangeAuthentication & Changes how API clients are authenticated: New, changed or deprecated API authentication methods\\
\hline
ChangeAuthorization & Changes who can access which API elements: Add, remove or modify permissions of API user\\
\hline
\end{tabular}
\end{table*}

@@

Another example that demonstrates a behavior change can be found from the 2013-10-29 version of Stripe API \footnote{\url{https://stripe.com/docs/upgrades\#2014-01-31}}:

@@

\textbf{Error condition change} - Web APIs change the error conditions that may require the users to adapt their integration. For example, in Salesforce API Version 29 the error codes were changed \footnote{\url{https://na1.salesforce.com/help/pdfs/en/salesforce_winter14_release_notes.pdf\#rn_186_api_objects}}:

@@

We analyzed different approaches to identify the versions of Web APIs to understand if the version identifiers can be used to infer information about their evolution. We can categorize the Web APIs into these groups:
\textbf{Numbered identifiers} - Contiguous integers are used to identify new versions by Salesforce and GitHub Web APIs. Such naming scheme clearly communicates the sequence between versions but cannot be used to infer compatibility between versions.

@@

\begin{figure*}[!hbtp]
  \begin{subfigure}[b]{\textwidth}
    \def\dataSourceFileName{facebookx.dat}
    \begin{tikzpicture}[show background rectangle,tight background]
      \begin{axis}[
          ybar stacked,
          bar width=5pt,
          date coordinates in=x,
          xticklabel = \year-\month,
          xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
          xlabel style={font=\small},
          ylabel style={font=\small},
          ylabel={No. of Changes},
          xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
          enlarge x limits=false,
          height=0.20\textheight,
          width=\textwidth,
          legend style={legend pos=north west,font=\tiny, legend columns=4},
          legend to name=legend
        ]
        \input{case_study/change_frequency}
      \end{axis}
    \end{tikzpicture}
    \label{fig:facebook-chart}
    \caption{Facebook API Changes (11 Releases)}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \def\dataSourceFileName{stripe.dat}
    \begin{tikzpicture}[show background rectangle,tight background]
      \begin{axis}[
          date coordinates in=x,
          ybar stacked,
          bar width=5pt,
          xticklabel = \year-\month,
          xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
    %xlabel={Timeline},
          xlabel style={font=\small},
          ylabel style={font=\small},
          ylabel={No. of Changes},
          xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
          enlarge x limits=false,
          height=0.20\textheight,
          width=\textwidth,
          legend to name=legend
        ]
        \input{case_study/change_frequency}
      \end{axis}
    \end{tikzpicture}
    \label{fig:stripe-chart}
    \caption{Stripe API Changes (30 Releases)}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \def\dataSourceFileName{wordpress.dat}
    \begin{tikzpicture}[show background rectangle,tight background]
      \begin{axis}[
    %title=WordPress REST API Changes (8 Releases),
          date coordinates in=x,
          ybar stacked,
          bar width=5pt,
          xticklabel = \year-\month,
          xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
    %xlabel={Timeline},
          xlabel style={font=\small},
          ylabel style={font=\small},
          ylabel={No. of Changes},
          xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
          enlarge x limits=false,
          height=0.20\textheight,
          width=\textwidth,
          legend to name=legend
        ]
        \input{case_study/change_frequency}
      \end{axis}
    \end{tikzpicture}
    \label{fig:wordpress-chart}
    \caption{WordPress REST API Changes (8 Releases)}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \def\dataSourceFileName{github.dat}
    \begin{tikzpicture}[show background rectangle,tight background]
      \begin{axis}[
    %title=GitHub API Changes (59 Releases),
          date coordinates in=x,
          ybar stacked,
          bar width=5pt,
          xticklabel = \year-\month,
          xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
    %xlabel={Timeline},
          xlabel style={font=\small},
          ylabel style={font=\small},
          ylabel={No. of Changes},
          xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
          enlarge x limits=false,
          height=0.20\textheight,
          width=\textwidth,
          legend to name=legend
        ]
        \input{case_study/change_frequency}
      \end{axis}
    \end{tikzpicture}
    \label{fig:github-chart}
    \caption{GitHub API Changes (59 Releases)}
  \end{subfigure}

\end{figure*}
\begin{figure*}[!h]\ContinuedFloat

  \begin{subfigure}[b]{\textwidth}
    \def\dataSourceFileName{gmap.dat}
    \begin{tikzpicture}[show background rectangle,tight background]
      \begin{axis}[
    %title=Google Maps Web API (102 Releases),
          date coordinates in=x,
          ybar stacked,
          bar width=5pt,
          xticklabel = \year-\month,
          xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
    %xlabel={Timeline},
          xlabel style={font=\small},
          ylabel style={font=\small},
          ylabel={No. of Changes},
          xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
          enlarge x limits=false,
          height=0.20\textheight,
          width=\textwidth,
          legend to name=legend
        ]
        \input{case_study/change_frequency}
      \end{axis}
    \end{tikzpicture}
    \label{fig:gmaps-chart}
    \caption{Google Maps Web API (102 Releases)}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \def\dataSourceName{Salesforce Force.com and Chatter REST}
    \def\dataSourceFileName{salesforce.dat}
    \begin{tikzpicture}[show background rectangle,tight background]
      \begin{axis}[
    %title=Salesforce REST API (4 Releases),
          date coordinates in=x,
          ybar stacked,
          bar width=5pt,
          xticklabel = \year-\month,
          xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
    %label={Timeline},
          xlabel style={font=\small},
          ylabel style={font=\small},
          ylabel={No. of Changes},
          xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
          enlarge x limits=false,
          height=0.20\textheight,
          width=\textwidth,
          legend to name=legend
        ]
        \input{case_study/change_frequency}
      \end{axis}
    \end{tikzpicture}
    \label{fig:salesforce-chart}
    \caption{Salesforce REST API (4 Releases)}
  \end{subfigure}

%% Add the legend here
  \begin{center}
    \begin{subfigure}[b]{10cm}
      \includegraphics[keepaspectratio, width=10cm]{legend.png}
      \caption{Legend}
    \end{subfigure}
  \end{center}
  \caption{Case Study - Change Profiles for the Web APIs.}
  \label{fig:proflie}
\end{figure*}

@@

Maintaining multiple versions of the Web APIs provide more scheduling flexibility to the API clients but requires more effort from the API publisher. As a result, not all the older versions are maintained. When API clients lag far behind the latest version, migrating to the latest version becomes harder. For example, here is an excerpt of a comment from a Stripe Web API client developer \footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/\#!searchin/api-discuss/version/api-discuss/x8TM-yHnhYQ/y9JJKytSiUAJ}}:

@@
